---
title: "R Notebook"
output: html_notebook
---


# Libraries
```{r}
library(dplyr) #data manipulation
library(tidyr) #data manipulation
library(psych) #descriptive statistics
library(Hmisc) #descriptive statistics

library(purrr)
library(tidyverse)
library(purrrlyr) #data type convertor

library(kableExtra)
library(xtable)

library(ggpubr) #multi plot
library(grid) #multi plot

library(caret) #machine learning
```

# Loading Data Into Environment
```{r}
winequality_white <- read_csv("Data/winequality-white.csv")
```

# Data Set Overview
```{r}
#You can as well use glimpse(), str() or summary() function
#The one below is a custom made one

head(winequality_white) %>%
  knitr::kable() %>%
    kable_styling(full_width = F, position = "left") %>%
    row_spec(0, color = "white", background = "#6897BB")
```

# Data Formating
```{r}
winequality_white <- winequality_white %>%
  select(quality, everything())

oldnames = c("fixed acidity","volatile acidity", "residual sugar", 
"free sulfur dioxide", "total sulfur dioxide", "citric acid")
newnames = c("fixedAcidity","volatileAcidity", "residualSugar", 
"freeSulfurdioxide", "totalSulfurDioxide", "citricAcid")

winequality_white <-winequality_white %>% 
  rename_at(vars(oldnames), ~ newnames) %>%
  mutate(quality = as.factor(ifelse(quality < 6, "low", "high")))

#You can do the same thing with the code below
colnames(winequality_white ) <- gsub(" ", "_", colnames(winequality_white ))
```


# Data Type Check Up
```{r}
#Select data types
data_types <- sapply(winequality_white, class)
data_types <- data.frame(feature = names(data_types), values = data_types, row.names = NULL)

#transpose the training set
values <- t(winequality_white)

#select the new transposed matrix and convert it to data frame class
datamatrix = as.data.frame(values, stringsAsFactors = FALSE)

#bind both data frames 
structure_df <- bind_cols(data_types, datamatrix)

#select only three columns 
structure_dfa <- structure_df[,1:3]

structure_dfa %>%
  knitr::kable() %>%
  kable_styling(full_width = F, position = "left") %>%
  row_spec(0, color = "white", background = "#6897BB", align = "center")
```

```{r}
winequality_white <- winequality_white %>%
  mutate_if(sapply(winequality_white, is.integer), as.factor)
```

```{r}
#the problem could be also solved like this:
#list the feature or features for which you would like to change data type
col <- c("quality")

#then use purrr's function dmap to change the data type of the feature
winequality_white[, col] <- winequality_white %>%
  select(one_of(col)) %>%
  dmap(as.factor)
```

# NAs Check
```{r}
#Sometimes values can be NA, but recorded under some other value. 
#Here are some of the potential NAs:
winequality_white[winequality_white == c("?", ".", ",", "!", "na")] <- NA

na <- winequality_white %>%
  map_df(function(x) {
    sum(is.na(x))
  }) %>%
  gather(Features, NA_values) %>%
  arrange(desc(NA_values)) %>%
  print(n = 15)
```
```{r}
na %>%
  knitr::kable() %>%
      kable_styling(full_width = F, position = "left") %>%
      row_spec(0, color = "white", background = "#6897BB")
```

# Spliting The Data

```{r}
set.seed(92)
trainIndex <- createDataPartition(winequality_white$quality, 
                                  p = .7,
                                  list = FALSE,
                                  times = 1)

train <- winequality_white[ trainIndex,]
test <- winequality_white[-trainIndex,]
```

# Visualize The Stratified Split

```{r}
rbind(data.frame(group = "train", train),
      data.frame(group = "test", test)) %>%
  gather(x, y, quality:sulphates) %>%
  ggplot(aes(x = y, color = group, fill = group)) +
    geom_density(alpha = 0.3) +
    facet_wrap( ~ x, scales = "free", ncol = 3)
```

```{r}
Hmisc::describe(train$quality)
```

```{r}
Hmisc::describe(test$quality)
```

# Exploratory Data Analysis
Descriptive Statistics
Numerical Data Type

```{r}
df_num <- train %>%
  select(-quality) %>%
  summarise_all(funs(min = min, 
                      Q25 = quantile(., 0.25), 
                      median = median, 
                      Q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd,
                      skewness = skew,
                      kurtosis = kurtosi))

# reshape it using tidyr functions
summary_num_stat <- df_num %>% 
  gather(stat, val) %>%
  separate(stat, into = c("features", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  select(features, min, Q25, Q75, max, median, mean, sd, skewness, kurtosis)

summary_num_stat %>%
  knitr::kable() %>%
    kable_styling(full_width = F, position = "left") %>%
    row_spec(0, color = "white", background = "#6897BB", align = "right")
```

# Categorical Data Type
```{r}
quality <- train %>%
  group_by(category) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))

quality %>%
  knitr::kable() %>%
    kable_styling(full_width = F, position = "left") %>%
    row_spec(0, color = "white", background = "#6897BB", align = "right")
```

# Univariate and Bivariate Relationships and Associations
Categorical Variables
```{r}
ggplot(data = train, aes(x = category)) +
   geom_bar(color = 'black', fill = "#6897BB") +
   geom_text(stat='count', aes(label=..count..), vjust=3, size=5, color = "white") +
   #theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
   guides(fill=FALSE) +
   xlab("quality") + ylab("Count") +
   labs(title="Count per category for 'quality' feature", 
   #subtitle="Count of All Positive and Negative Samples", 
   caption="Source: winequality_white")
```

Continuous Variables

```{r}
train %>%
  gather(-quality, key = "var", value = "value") %>% 
  ggplot(aes(x = value, fill = as.factor(quality))) +
  geom_histogram(color = "white", bins = 30, alpha = 0.6) +
  facet_wrap(~ var, scales = "free") +
  ylab("Frequency") +
  xlab("All Continuous Features") +
  labs(title = "The distribution of low and high quality wine",
       subtitle = "Average") +
  #scale_fill_brewer(name = "New Legend Title") +
  scale_fill_manual(values = c("#ff1919", "#0e2b58"), name = "Quality") +
  theme(text = element_text(color = "gray20"),
        legend.position = c("right"), # position the legend in the upper left 
        legend.direction = "vertical",
        legend.justification = 0.1, # anchor point for legend.position.
        legend.text = element_text(size = 11, color = "gray10"),
        axis.text = element_text(face = "italic"),
        axis.ticks.y = element_blank(), # element_blank() is how we remove elements
        axis.line.y = element_blank(),
        panel.grid.major = element_line(),
        panel.grid.major.x = element_blank())
```

```{r}
train %>%
  gather(-quality, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y= value, fill = as.factor(quality))) +
  geom_boxplot() +
  #scale_y_log10() +
  #scale_x_log10() +
  facet_wrap(~ var, scales = "free") +
  ylab("Feature") +
  xlab("Feature") +
  labs(title = "'Box Plot per Quality'") +
  #scale_fill_manual(name = "Boxplot", palette = "Paired") +
  scale_fill_manual(values = c("#ff1919", "#0e2b58"), name = "Quality") +
  theme(text = element_text(color = "gray20"),
        legend.position = c("right"), # position the legend in the upper left 
        legend.direction = "vertical",
        legend.justification = 0.1, # anchor point for legend.position.
        legend.text = element_text(size = 11, color = "gray10"),
        axis.text.x=element_blank(), #hides x axis
        axis.text = element_text(face = "italic"),
        axis.ticks.y = element_blank(), # element_blank() is how we remove elements
        axis.line.y = element_blank(),
        panel.grid.major = element_line(),
        panel.grid.major.x = element_blank())
```

```{r}
train %>%
  gather(-quality, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y= value, fill = as.factor(quality))) +
  geom_violin(alpha = 0.6) +
  #scale_y_log10() +
  #scale_x_log10() +
  facet_wrap(~ var, scales = "free") +
  ylab("Features") +
  xlab("Features") +
  labs(title = "'Violin Plot per Quality'") +
  #scale_fill_brewer(name = "Boxplot", palette = "Paired")
  #scale_fill_manual(name = "Boxplot", palette = "Paired") +
  scale_fill_manual(values = c("#ff1919", "#0e2b58"), name = "Quality") +
  theme(text = element_text(color = "gray20"),
        legend.position = c("right"), # position the legend in the upper left 
        legend.direction = "vertical",
        legend.justification = 0.1, # anchor point for legend.position.
        legend.text = element_text(size = 11, color = "gray10"),
        axis.text = element_text(face = "italic"),
        axis.ticks.y = element_blank(), # element_blank() is how we remove elements
        axis.line.y = element_blank(),
        panel.grid.major = element_line(),
        panel.grid.major.x = element_blank())
```

```{r}
train %>%
  gather(-quality, -alcohol, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = alcohol, color = quality)) +
  geom_point(alpha = 0.25) +
    scale_color_manual(name = "",
                     values = c("#ff1919", "#0e2b58")) +
  facet_wrap(~ var, scales = "free") + 
  labs(title = "Scatterplot") +
  theme(text = element_text(size=9, family = "Georgia"),
        axis.ticks = element_blank(),
        plot.title = element_text(size = 15, margin = margin(b = 8)),
        plot.subtitle = element_text(size = 10, color = "darkslategrey", margin = margin(b = 20)),
        plot.caption = element_text(size = 6, margin = margin(t = 8), color = "grey70", hjust = 0),
        strip.text = element_text(size = 10, face="bold"))
```


```{r}
class_comparison <- train %>%
  #Creating new feature by breaking alcohol into 2 levels 
  mutate(category=cut(alcohol, breaks=c(-Inf, 10, Inf), labels=c("low_alcohol", "high_alcohol"))) %>%
  select(quality, category, alcohol) 

#being more explicit in regards to what is meant by low and high
levels(class_comparison$quality) <- c('high_quality', 'low_quality')

#Calculating mean for each category
class_mean <- class_comparison %>%
  group_by(quality, category) %>%
  summarise(Mean = mean(alcohol))

#Creaating label and rounding it to max 3 decimals. 
class_mean <- class_mean %>%
  mutate(Label = paste0(round(Mean, 3)))

#Histogram Plot
ggplot(class_comparison, aes(x = alcohol, fill = as.factor(category))) +
  geom_histogram(bins = 30, color = "white") +
  scale_fill_manual(values=c("#F8766D", "#6897BB")) +
  geom_vline(data = class_mean, aes(xintercept = Mean), linetype = "dashed", alpha = .6) +
  facet_grid(quality ~ category, scales = "free") +
  scale_y_continuous(breaks=pretty_breaks(n=2), limits = c(0, 150)) +
  ylab("Frequency") +
  xlab("Alcohol") +
  labs(title = "Mean Alcohol value devided by Quality & Alcohol - low/high levels") +
  #subtitle = "The higher the level of wine, the greater is chance that it is considered high quality") +
  geom_text(data = class_mean, aes(x = Mean, y = 130, id = quality, label = Label),
            family = "Georgia", size = 3, hjust = -.1) +
  theme(text = element_text(size=9, family = "Georgia"),
        legend.position='none',
        axis.ticks = element_blank(),
        plot.title = element_text(size = 15, margin = margin(b = 8)),
        plot.subtitle = element_text(size = 10, color = "darkslategrey", margin = margin(b = 20)),
        plot.caption = element_text(size = 6, margin = margin(t = 8), color = "grey70", hjust = 0),
        strip.text = element_text(size = 10, face="bold"))
```

Correlation Matrix
```{r}
#First thing change the data type from character to integer, so that cor function can work
train_a <- train %>%
  mutate_if(sapply(train, is.factor), as.integer)

train_a$quality <- ifelse(train_a$quality == 2, 0, train_a$quality) 

#implement cor function
M <- cor(train_a)
#head(round(M,2))

# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat<- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(train_a)

#Correlation plot
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         tl.cex = 0.75, 
         number.cex = .55,
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE)
```

```{r}
res2 <- rcorr(as.matrix(train))
cor_sig <- as.data.frame(res2$P)

cor_sig[is.na(cor_sig)] <- 0

cor_sig[,1:12] <- round(cor_sig, digits = 3)

cor_sig <- cor_sig %>% 
  mutate(
  features = row.names(.),
  quality = cell_spec(quality, color = ifelse(quality > 0.05, "red", "black")),
  fixedAcidity = cell_spec(fixedAcidity, color = ifelse(fixedAcidity > 0.05, "red", "black")),
  volatileAcidity = cell_spec(volatileAcidity, 
color = ifelse(volatileAcidity > 0.05, "red", "black")),
  citricAcid = cell_spec(citricAcid, color = ifelse(citricAcid > 0.05, "red", "black")),
  residualSugar = cell_spec(residualSugar, color = ifelse(residualSugar > 0.05, "red", "black")),
  chlorides = cell_spec(chlorides, color = ifelse(chlorides > 0.05, "red", "black")),
  freeSulfurdioxide = cell_spec(freeSulfurdioxide, 
color = ifelse(freeSulfurdioxide > 0.05, "red", "black")),
  totalSulfurDioxide = cell_spec(totalSulfurDioxide, 
color = ifelse(totalSulfurDioxide > 0.05, "red", "black")),
  density = cell_spec(density, color = ifelse(density > 0.05, "red", "black")),
  pH = cell_spec(pH, color = ifelse(pH > 0.05, "red", "black")),
  sulphates = cell_spec(sulphates, color = ifelse(sulphates > 0.05, "red", "black")),
  alcohol = cell_spec(alcohol, color = ifelse(alcohol > 0.05, "red", "black")))


cor_sig %>%
  #select_all() %>%
  kable(escape = F) %>%
  kable_styling("striped", full_width = F) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right") %>%
  row_spec(0, color = "white", background = "#6897BB", align = "center")
```

Feature Engineering

```{r}
trainFE <- train %>%
  mutate(
    
    #Isolation features
    chloridesHIGH = chlorides < 0.06,
    citricAcidBETWEEM = citricAcid < 0.5 & citricAcid > 0.1,
    freeSulfurdioxideHIGH = freeSulfurdioxide > 65,
    pHHIGH = pH > 3.40,
    residualSugarHIGH = residualSugar > 14,
    sulphatesHIGH = sulphates > 0.6,
    volatileAcidityHIGH = volatileAcidity > 0.45,
    sulphatesHIGH = sulphates > 0.6,
    fixedAcidityLOW = fixedAcidity < 6,
    alcoholLOW = alcohol < 9,
    alcoholHIGH = alcohol > 12.5,
    
    #Interaction features
    chloridesAL = chlorides * alcohol,
    fixedAcidityAL = fixedAcidity * alcohol,
    volatileAcidityAL = volatileAcidity * alcohol,
    citricAcidAL = citricAcid * alcohol,
    residualSugarAL = residualSugar * alcohol,
    chloridesAL = chlorides * alcohol, 
    freeSulfurdioxideAL = freeSulfurdioxide * alcohol,
    totalSulfurDioxideAL = totalSulfurDioxide * alcohol,
    densityAL = density * alcohol,
    pHAL = pH * alcohol
  )
```


```{r}
col <- c("chloridesHIGH", "citricAcidBETWEEM", "freeSulfurdioxideHIGH", "pHHIGH", 
"residualSugarHIGH", "sulphatesHIGH", "volatileAcidityHIGH", "fixedAcidityLOW", 
"alcoholLOW", "alcoholHIGH")

#take columns from col object and transform it from logical data type to a factor
trainFE[, col] <- trainFE %>%
  select(one_of(col)) %>%
  dmap(as.factor)

head(x) %>%
  knitr::kable() %>%
    kable_styling(full_width = F, position = "left") %>%
    row_spec(0, color = "white", background = "#6897BB", align = "right")
```

Modeling

```{r}
trControl = trainControl(method = "repeatedcv", 
                         number = 5, 
                         repeats = 5, 
                         savePredictions = TRUE, 
                         verboseIter = FALSE)
```

```{r}
set.seed(92)
model_lm <- caret::train(quality ~ .,
                         data = train,
                         method = "lm",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)

model_lmFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "lm",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)

model_glm <- caret::train(quality ~ .,
                         data = train,
                         method = "glm",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)

model_glmFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "glm",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)


library(elasticnet)
model_lasso <- caret::train(quality ~ .,
                         data = train,
                         method = "lasso",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

library(elasticnet)
model_lassoFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "lasso",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

model_ridge <- caret::train(quality ~ .,
                         data = train,
                         method = "ridge",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

model_ridgeFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "ridge",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

library(glmnet)
model_glmnet <- caret::train(quality ~ .,
                         data = train,
                         method = "glmnet",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)



library(glmnet)
model_glmnetFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "glmnet",
                         preProcess = c("BoxCox","scale", "center"),
                         trControl = trControl)

model_rf <- caret::train(quality ~ .,
                         data = train,
                         method = "rf",
                         preProcess = c("scale", "center"),
                         trControl = trControl)

model_rfFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "rf",
                         preProcess = c("scale", "center"),
                         trControl = trControl)

set.seed(92)
model_Cart <- caret::train(quality ~ .,
                         data = train,
                         method = "rpart",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

model_CartFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "rpart",
                         preProcess = c("BoxCox", "scale", "center"),
                         trControl = trControl)

model_knn <- caret::train(quality ~ .,
                         data = train,
                         method = "knn",
                         preProcess = c("scale", "center"),
                         trControl = trControl)

model_knnFE <- caret::train(quality ~ .,
                         data = trainFE,
                         method = "knn",
                         preProcess = c("scale", "center"),
                         trControl = trControl)
```


```{r}
# Compare model performances using resample()
models_compare <- resamples(list(LM = model_lm, LM_FE = model_lmFE, 
                                 GLM = model_glm, GLM_FE = model_glmFE,
                                 Lasso = model_lasso, LassoFE = model_lassoFE, 
                                 Ridge = model_ridge, RidgeFE = model_ridgeFE, 
                                 Elastic_Net = model_glmnet, Elastic_NetFE = model_glmnetFE,
                                 RF = model_rf, RFFE = model_rfFE,
                                 CART = model_Cart, CARTFE = model_CartFE, 
                                 KNN = model_knn, KNNFE = model_knnFE))

# Summary of the models performances
summary(models_compare)
```

```{r}
dotplot(models_compare)
```

```{r}
bwplot(models_compare)
```

```{r}
# dot plots of accuracy
scales <- list(x=list(relation="free"), y=list(relation="free"))
dotplot(models_compare, scales=scales)
```

```{r}
parallelplot(models_compare)
```

```{r}
xyplot(models_compare, models=c("RF", "RFFE"))
```

```{r}
varimp_mars <- varImp(model_glmnet)
plot(varimp_mars, main="Variable Importance with RF")
```








